Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-02-05 ms.tex
Adding table which shows cooling solution.

\begin{table*}
\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions}
\begin{tabular}{lccccccccccccc}
\hline
{$M_{\bullet,8}$} & {$v_{w}$} & {$\Gamma^{(a)}$} & {$r_b$} & {$\dot{M}/\dot{M}_{\rm edd}$} & {$(t_{\rm cool}/t_{\rm ff})_{r_s}$} & {comments}\\
\hline
0.01 & 300 & 0.8 & U & $ 3.2 \times 10^{ -4 }$ & 4.7 & rb in fact
unresolved \\
0.01 & 600 & 0.8 & U & $ 5.5 \times 10^{ -5 }$ & $ 3.2 \times
10^{ 2 }$ &  \\
0.01 & 1200 & 0.8 & U & $ 9.4 \times 10^{ -6 }$ & $ 1.5 \times
10^{ 4 }$ &  \\
0.1 & 300 & 0.8 & 100.0 & $ 2.5 \times 10^{ -3 }$ & 0.29 & \\
0.1 & 600 & 0.8 & 100.0 & $ 1.8 \times 10^{ -4 }$ & 97 &  \\
0.1 & 1200 & 0.8 & U & $ 3.1 \times 10^{ -5 }$ & $ 4.8 \times
10^{ 3 }$ &  \\
1 & 300 & 0.8 & 100.0 & $ 1.6 \times 10^{ -2 }$ & $ 5.6 \times 10^{
  -2 }$ &  TI\\
1 & 600 & 0.8 & 100.0 & $ 7.3 \times 10^{ -4 }$ & 21 &  \\
1 & 1200 & 0.8 & 100.0 & $ 9.9 \times 10^{ -5 }$ & $ 1.5 \times 10^{
  3 }$ &  \\
1 & 300 & 0.8 & 50.0 & $ 8.2 \times 10^{ -3 }$ & 0.19 &  \\
1 & 300 & 0.8 & 200.0 & $ 3.6 \times 10^{ -2 }$ & $ 1.7 \times 10^{
  -2 }$ &  TI/convergence \\
%1 & 300.0 & 0.8 & 300.0 & $ 5.2 \times 10^{ -2 }$ & $ 9.8 \times 10^{
%  -3 }$ &  TI/convergence\\
1 & 300 & 0.8 & 400.0 &  &  &  TI/run unsuccessful\\
0.1 & 300 & 0.8 & 50.0 & $ 2.0 \times 10^{ -3 }$ & 0.46 &  \\
0.1 & 300 & 0.8 & 200.0 & $ 5.3 \times 10^{ -3 }$ & 0.12 &  TI/convergence\\
%0.1 & 300.0 & 0.8 & 300.0 & $ 7.2 \times 10^{ -3 }$ & $ 6.7 \times
%10^{ -2 }$   & TI/convergence  \\
0.1 & 300 & 0.8 & 400.0 & $ 8.9 \times 10^{ -3 }$ & $ 4.7 \times
10^{ -2 }$ & TI/convergence \\
0.01 & 1200 & 0.1 & U & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times
10^{ 4 }$ &  \\
0.1 & 1200 & 0.1 & U & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{
  3 }$ &  tried run in which rb=100 pc was resolved what would happen if rb=400 pc?\\
0.01 & 1200 & 0.1 & 100.0 & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times 10^{ 4 }$ &  \\
0.1 & 1200& 0.1 & 100.0 & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 100.0 & $ 6.0 \times 10^{ -5 }$ & $ 8.6 \times 10^{ 2 }$ &  \\
1 & 1200 & 0.1 & 50.0 & $ 4.9 \times 10^{ -5 }$ & $ 1.2 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 200.0 & $ 1.0 \times 10^{ -4 }$ & $ 3.4 \times 10^{ 2 }$ &  \\
%1 & 1200.0 & 0.1 & 300.0 & $ 1.9 \times 10^{ -4 }$ & $ 1.3 \times 10^{ 2 }$ &  convergence\\
1 & 1200 & 0.1 & 400.0 & $ 1.8 \times 10^{ -2 }$ & 0.37 &  TI/convergence\\
\hline
\label{table:models}  
\end{tabular}
$^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$
Break radius of stellar Nuker profile.  For profiles marked U, the
break radius is not resolved on the grid.
\end{minipage}
\end{table*}

@@

{\bf AG: Note that the Compton heating will depend on the distance of
  the hole as $1/r^2$ so Compton heating will be stronger closer to the
  black hole! }
The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective Compton heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 29 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 50 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  If heating is provided by
Compton heating alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then equation
(\ref{eq:vC}) can be solved for $\tilde{v}_{\rm w}$ to find
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 106 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 76 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
The fact that $v_{w}^{\rm C} \lesssim 200$ km s$^{-1}$ for physically realized values of $\eta \lesssim 1$ shows that Compton heating alone is insufficient to stabilize the flow against cooling
instability, i.e. $v_{w}^{C} < v_{\rm CI}$ (eq.~[\ref{eq:cooling2}]).

Kinetic feedback occurs from an outflow of mass or energy  from the viscinity of the black hole in the form of a disk wind or jet, which deposits its energy across a much wider range of radii.  If the total outflow power is proportional to the SMBH accretion rate, $L_{\rm j} = \epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is an efficiency factor.  Assuming that this energy is deposited uniformly out to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm heat}^{3}$, then the volumetric heating rate $e = \epsilon_{\rm j}\dot{M}c^{2}/V_{\rm heat}$ results in an effective wind velocity at the stagnation radius given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray} where we have used the fact that $\dot{M} \approx \eta
M_{\rm enc}(r_{\rm s})/t_{\rm h}$ and that $M_{\rm enc}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ for radii interior to the outer
break radius (eq.~[\ref{eq:rhostar}]).  Thus for $r_{\rm heat} \sim
r_{\rm s}$, even a modest heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other
sources of non-accretion powered heating.  However, if the jet distributes its energy on much larger scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ kpc $\gtrsim 10^{3}r_{\rm s}$ then the heating rate would be much less significant, even for an efficient jet ($\epsilon_{\rm j} \sim 1$).

@@

What is a physically motivated value of the heating radius $r_{\rm
  heat}$?  \citet{Bromberg+11} show that a jet of luminosity $L_{\rm
  j}$ and half opening angle $\theta_{\rm j} = 0.1$ (characteristic of
AGN jets) propagates through a gaseous mass $M_{\rm g}$ to a radius
$r$ in a timescale
\begin{eqnarray}
t_{\rm jet} &\sim& 4000\,{\rm yr}\left(\frac{{L}_{\rm j}}{10^{40}\,\rm erg\,s^{-1}}\right)^{-1/3}\left(\frac{r}{\rm pc}\right)^{2/3}\left(\frac{M_{\rm g}}{10^{8}M_{\odot}}\right)^{1/3} 
\end{eqnarray}
Writing $L_{\rm j} = \epsilon_{\rm j}\dot{M}c^{2}$ and approximating $M_{\rm g} \sim \dot{M}t_{\rm ff}$ we find that the ratio of the jet crossing time to the free-fall time $t_{\rm ff} \sim r/\sigma$ is 
\begin{equation}
\frac{t_{\rm jet}}{r/\sigma} \sim 8\times 10^{-5}\sigma_{200}^{2/3}\epsilon_{\rm j}^{-1/3},
\end{equation}
which is notably independent of radius $r$.  Thus we require $\epsilon_{\rm j} \gtrsim 10^{-12}$ for the jet to escape, and if it escapes it is free to travel a long distance.

>>>>

\begin{table*}
\centering
\begin{minipage}{18cm}
\caption{Summary of Numerical Solutions (without cooling or conductivity)}
\begin{tabular}{lccccccccccccc}
\hline
{$M_{\bullet,8}$} & {$v_{w}$} & {$\Gamma^{(a)}$} & {$r_b$} &
{$\dot{M}/\dot{M}_{\rm edd} \eta$} & {$(t_{\rm cool}/t_{\rm ff})_{r_s} \eta^{-1}$} & {comments}\\
\hline
0.01 & 300 & 0.8 & U & $ 3.2 \times 10^{ -4 }$ & 4.7 & rb in fact
unresolved \\
0.01 & 600 & 0.8 & U & $ 5.5 \times 10^{ -5 }$ & $ 3.2 \times
10^{ 2 }$ &  \\
0.01 & 1200 & 0.8 & U & $ 9.4 \times 10^{ -6 }$ & $ 1.5 \times
10^{ 4 }$ &  \\
0.1 & 300 & 0.8 & 100.0 & $ 2.5 \times 10^{ -3 }$ & 0.29 & \\
0.1 & 600 & 0.8 & 100.0 & $ 1.8 \times 10^{ -4 }$ & 97 &  \\
0.1 & 1200 & 0.8 & U & $ 3.1 \times 10^{ -5 }$ & $ 4.8 \times
10^{ 3 }$ &  \\
1 & 300 & 0.8 & 100.0 & $ 1.6 \times 10^{ -2 }$ & $ 5.6 \times 10^{
  -2 }$ &  TI\\
1 & 600 & 0.8 & 100.0 & $ 7.3 \times 10^{ -4 }$ & 21 &  \\
1 & 1200 & 0.8 & 100.0 & $ 9.9 \times 10^{ -5 }$ & $ 1.5 \times 10^{
  3 }$ &  \\
1 & 300 & 0.8 & 50.0 & $ 8.2 \times 10^{ -3 }$ & 0.19 &  \\
1 & 300 & 0.8 & 200.0 & $ 3.6 \times 10^{ -2 }$ & $ 1.7 \times 10^{
  -2 }$ &  TI/convergence \\
%1 & 300.0 & 0.8 & 300.0 & $ 5.2 \times 10^{ -2 }$ & $ 9.8 \times 10^{
%  -3 }$ &  TI/convergence\\
1 & 300 & 0.8 & 400.0 &  &  &  TI/run unsuccessful\\
0.1 & 300 & 0.8 & 50.0 & $ 2.0 \times 10^{ -3 }$ & 0.46 &  \\
0.1 & 300 & 0.8 & 200.0 & $ 5.3 \times 10^{ -3 }$ & 0.12 &  TI/convergence\\
%0.1 & 300.0 & 0.8 & 300.0 & $ 7.2 \times 10^{ -3 }$ & $ 6.7 \times
%10^{ -2 }$   & TI/convergence  \\
0.1 & 300 & 0.8 & 400.0 & $ 8.9 \times 10^{ -3 }$ & $ 4.7 \times
10^{ -2 }$ & TI/convergence \\
0.01 & 1200 & 0.1 & U & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times
10^{ 4 }$ &  \\
0.1 & 1200 & 0.1 & U & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{
  3 }$ &  tried run in which rb=100 pc was resolved what would happen if rb=400 pc?\\
0.01 & 1200 & 0.1 & 100.0 & $ 1.5 \times 10^{ -6 }$ & $ 4.1 \times 10^{ 4 }$ &  \\
0.1 & 1200& 0.1 & 100.0 & $ 9.5 \times 10^{ -6 }$ & $ 5.7 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 100.0 & $ 6.0 \times 10^{ -5 }$ & $ 8.6 \times 10^{ 2 }$ &  \\
1 & 1200 & 0.1 & 50.0 & $ 4.9 \times 10^{ -5 }$ & $ 1.2 \times 10^{ 3 }$ &  \\
1 & 1200 & 0.1 & 200.0 & $ 1.0 \times 10^{ -4 }$ & $ 3.4 \times 10^{ 2 }$ &  \\
%1 & 1200.0 & 0.1 & 300.0 & $ 1.9 \times 10^{ -4 }$ & $ 1.3 \times 10^{ 2 }$ &  convergence\\
1 & 1200 & 0.1 & 400.0 & $ 1.8 \times 10^{ -2 }$ & 0.37 &  TI/convergence\\
\hline
\label{table:models}  
\end{tabular}
$^{(a)}$ Inner power-law index of stellar Nuker profile.  $^{(b)}$
Break radius of stellar Nuker profile.  For profiles marked U, the
break radius is not resolved on the grid.
\end{minipage}
\end{table*}

\begin{table*}
\centering
\begin{minipage}{18cm}
\caption{Comparison of numerical solutions with and without
  cooling. We have chosen $\eta=0.04$, which one would expect for 
  the average star formation history for $\Mbh=10^8\Msun$ 
  (see Figures~\ref{fig:eta} {\bf AG-cite star formation history
    figure once it is made.}). Note that we have deliberately chosen a solution
  near our thermal stability threshold where cooling is beginning
  to become important.
}
\begin{tabular}{lccccccccccccc}
\hline
{$M_{\bullet,8}$} & {$v_{w}$} & {$\Gamma$} & {$r_b$} & {$\eta$} &
{$\dot{M}/\dot{M}_{\rm edd}$} & {$(t_{\rm cool}/t_{\rm ff})_{r_s}$} & {comments}\\
\hline
1 & 300 & 0.8 & 50.0 & 0.04 & $ 3.3 \times 10^{ -4 }$ & 4.6 &  \\
1 & 300 & 0.8 & 50.0 & 0.04 & $ 9.1 \times 10^{ -4 }$ & 5.6 & \\
\hline
\label{table:models}  
\end{tabular}
\end{minipage}
\end{table*}

@@

{\bf AG: Note that the Compton heating will depend on the distance of
  the hole as $1/r^2$ so Compton heating will be stronger closer to the
  black hole! }
The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective Compton heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 29 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 50 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  If heating is provided by
Compton heating alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then equation
(\ref{eq:vC}) can be solved for $\tilde{v}_{\rm w}$ to find

\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 106 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 76 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align}

The fact that $v_{w}^{\rm C} \lesssim 200$ km s$^{-1}$ for physically
realized values of $\eta \lesssim 1$ shows that Compton heating alone
is insufficient to stabilize the flow against cooling instability,
i.e. $v_{w}^{C} < v_{\rm CI}$ (eq.~[\ref{eq:cooling2}]).

Kinetic feedback occurs from an outflow of mass or energy from the
viscinity of the black hole in the form of a disk wind or jet, which
deposits its energy across a much wider range of radii.  If the total
outflow power is proportional to the SMBH accretion rate, $L_{\rm j} =
\epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is an
efficiency factor.  Assuming that this energy is deposited uniformly
out to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm
heat}^{3}$, then the volumetric heating rate $e = \epsilon_{\rm
j}\dot{M}c^{2}/V_{\rm heat}$ results in an effective wind velocity at
the stagnation radius given by

\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray}

where we have used the fact that $\dot{M} \approx \eta M_{\rm
enc}(r_{\rm s})/t_{\rm h}$ and that $M_{\rm enc}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ for radii interior to the outer
break radius (eq.~[\ref{eq:rhostar}]).  Thus for $r_{\rm heat} \sim
r_{\rm s}$, even a modest heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other sources
of non-accretion powered heating.  However, if the jet distributes its
energy on much larger scales comparable to the size of the galaxy,
i.e. $r_{\rm heat} \gtrsim $ kpc $\gtrsim 10^{3}r_{\rm s}$ then the
heating rate would be much less significant, even for an efficient jet
($\epsilon_{\rm j} \sim 1$).

@@

What is a physically motivated value of the heating radius $r_{\rm
  heat}$?  \citet{Bromberg+11} show that a jet of luminosity $L_{\rm
  j}$ and half opening angle $\theta_{\rm j} = 0.1$ (characteristic of
AGN jets) propagates through a gaseous mass $M_{\rm g}$ to a radius
$r$ in a timescale

\begin{eqnarray}
t_{\rm jet} &\sim& 4000\,{\rm yr}\left(\frac{{L}_{\rm j}}{10^{40}\,\rm erg\,s^{-1}}\right)^{-1/3}\left(\frac{r}{\rm pc}\right)^{2/3}\left(\frac{M_{\rm g}}{10^{8}M_{\odot}}\right)^{1/3} 
\end{eqnarray}

Writing $L_{\rm j} = \epsilon_{\rm j}\dot{M}c^{2}$ and approximating
$M_{\rm g} \sim \dot{M}t_{\rm ff}$ we find that the ratio of the jet
crossing time to the free-fall time $t_{\rm ff} \sim r/\sigma$ is

\begin{equation}
\frac{t_{\rm jet}}{r/\sigma} \sim 8\times 10^{-5}\sigma_{200}^{2/3}\epsilon_{\rm j}^{-1/3},
\end{equation}

which is notably independent of radius $r$.  Thus we require
$\epsilon_{\rm j} \gtrsim 10^{-12}$ for the jet to escape, and if it
escapes it is free to travel a long distance.
